\textcolor{red}{A Software Tool for building Suitable Datasets for Softcomputing and Machine Learning Techniques from Meteorological Data Integration}

### 2020-09-15 latex_manuscript/MDPI/2_energies_special_issue/template.tex
Cambios para el Special Issue de Energies

%=================================================================
% Full title of the paper (Capitalized)
\Title{Nuevo titulo relacionado con el Special Issue. El actual es SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets}

@@

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{Meteorological data are extensively used to perform environmental learning. Machine Learning (ML) techniques, a valuable support in many research areas, require datasets containing information related to the topic under study, which are not always available in an appropriate format, and its preparation and pre-processing implies a lot of time and effort by the researchers. This paper presents a novel software tool with an user-friendly GUI to create datasets by means of management and data integration of meteorological observations from two well-known sources of information: the \textit{National Data Buoy Center} and the \textit{National Centers for Environmental Prediction and for Atmospheric Research Reanalysis Project}. Such datasets can be created using buoys and reanalysis data by customisable procedures, in terms of predictive and objective variables and temporal resolution. These datasets can be used by ML methodologies for prediction tasks (classification or regression) that support improvement of sustainability energy production, design of production systems as WECs or environmental modelling, among others.}

% Keywords
\keyword{Meteorological data; Reanalysis data; Creating datasets; Environmental prediction; Marine energy; AQUI TAMBIEN HABRA QUE VER QUE PONEMOS ???????????}  % List three to ten pertinent keywords specific to the article, yet reasonably common within the subject discipline.

@@

		A better understanding of the environment is of vital importance for science, contributing not only to more efficient exploitation of natural resources but also to the development of new strategies aimed at its protection. In that sense, meteorological observations provide an essential and valuable source of information which is widely used by researchers to address environmental learning, comprehension, prediction and conservation in numerous oceanic and atmospheric studies of a wide variety of areas (e.g. energy, climate change, agriculture, etc.). Some specific examples of the diversity of fields in which meteorological data can be used in are, among others: estimation of global solar radiation based on sunshine duration \cite{SHAHRUKHANIS2019179}, directional analysis of sea storms \cite{LAFACE201545}, assessment and control of wind turbine \cite{DAHHANI2018167}, wind power ramp events prediction \cite{DORADOMORENO2017428}, evaluation of salinity-gradient energy harvesting \cite{REYESMENDOZA2020100645}, study of the responses exhibited by plankton to fluid motions \cite{FUCHS2016109}, trends in solar radiation \cite{SILVA20101852} or simulation of extreme near shore sea conditions \cite{GOULDBY201415}. All these studies require a prior data collection and its adaptation to a specific format that allows the interpretation of them.

		On the other hand, special purpose software is usually developed to help researchers to advance in their studies related to energy and environmental modelling, becoming a great support for decision-making in the exploitation and protection of the environment. In \cite{CAMARGONOGUEIRA2016361}, a software tool for designing solar water heating systems is developed, which simulates different situations and finds the best technical and economical solution. In \cite{LO2015293}, an integrated simulation tool for the optimum design of bifacial solar panel with reflectors is presented. This tool can also be used to analyse the performance of the solar cells. A framework for data integration of offshore wind farms is implemented in \cite{NGUYEN2013150} in order to facilitate data exchange and improve operation and maintenance practices. In \cite{SHEHADEH2020100644}, a load management application to maximize the used energy of thermal systems and reduce the costs and emissions is developed. A risk assessment tool to improve safety standards and emergency management in onshore wind farms, is presented in \cite{ASTIASOGARCIA201648}. Raabe et al. \cite{RAABE2010213} developed two software tools, \textit{Model of Equilibrium of Bay Beaches} (MEPBAY) and \textit{Coastal Modelling System} (SMC), to support different operational levels of headland-bay beach in coastal engineering projects, and Motahhir et al. \cite{MOTAHHIR20199} developed an open hardware/software test bench for solar tracker.

@@

		Currently, and as a support to traditional study procedures, Machine Learning (ML) techniques \cite{Alpaydin:2004:IML:1036287,Bishop:2006:PRM:1162264} are being widely used in numerous research fields related to classification, regression and optimisation tasks, obtaining significant improvements in the performance of the results \cite{DELFANI2019100559,NASRUDDIN201948,GUO201816}. ML methodologies can be used not only by experienced computer scientists but also by other researchers. For example, the well-known \textit{Waikato Environment for Knowledge Analysis} (WEKA) \cite{WEKA} software tool provides researchers with a wide collection of ML algorithms. In this way, ML techniques have been already applied to tackle wave characterisation, accurately estimating $H_{s}$ and $T_{e}$ parameters \cite{DURANROSAL2017268, KUMAR2017605}, given that robustness of ML methods can tackle the previously explained difficulties in wave energy prediction. The problem is that, in order to apply these methods, it is essential to obtain datasets with relevant information about the issue under study, used to infer knowledge. Usually, these datasets are not publicly available in a friendly format, and their generation is the first step needed.

@@

		Up to now, and to the best of our knowledge, this is the first software tool addressing the problem previously discussed, integrating meteorological data from NDBC and NNRP. The novelties and functionalities that SPAMDA offers to the researchers will be detailed in Section \ref{sec:SPAMDA}, although some of them are briefly summarised below:
		\begin{itemize}
			\item The generation of datasets becomes a very easy and customisable task, by means of the selection of different input parameters, such as predictive and objective variables, classification and regression ML techniques, output discretisation or prediction horizon, among others.
			\item It makes the researcher focus on environmental modelling, without having to worry about mechanical tasks, avoiding laborious pre-processing procedures that imply a lot of time and effort.
			\item It manages the extensive casuistry of data integration which can lead to incomplete datasets.
			\item It avoids possible researcher errors in the intermediate steps of the process, such as time and geographical coordinates conversion or missing values handling (dates or measurements not recorded), among others.
			\item It provides information about the quality and quantity of the data.
			\item It includes different pre-processing tasks, such as normalisation and missing data recovery.
			\item It facilitates data management and well-organised storage of the datasets.
			\item The created datasets can be easily used by ML tools.
			\item Its modular design allows the implementation of new modules for managing meteorological data from others sources, benefiting future renewable energy and environmental research.
			\item It includes a user-friendly GUI, facilitating and greatly simplifying data management, and it is integrated with the Explorer environment of WEKA.
			\item It is multi-platform, and it can be used on any computer with Java regardless of the operating system.
		\end{itemize}

@@

			Note that the data collected by the network of buoys may be incomplete due to diverse circumstances such as the weather conditions in which the buoys have to operate, failures or malfunctioning elements of the buoys, among others. Accordingly, it may be the situation that some of the measurements are completely missing (missing date or instance) or partially missing (some measurements not recorded), by a buoy or by a set of buoys, once in a while or over a period of time. It may be also possible that the measurements have been recorded at a time different from the expected one. These aspects have to be taken into account when creating the datasets. This casuistry is explained in detail in Appendix A \ref{app:AppendixA}.

@@

@@

				The reader can check in Appendix A \ref{app:AppendixA} an example with a more complex case of the procedure.
				\begin{figure}[ht!]
					\centering
					\includegraphics[scale=0.38]{figures/FigureMatchingProcess.png}
					\caption{An example of matching the data from NDBC (left) and NNRP (right).}
					\label{fig:matchingProcess}
				\end{figure}

@@

\appendix
\section{Managing incomplete data}\label{app:AppendixA}
		In this appendix, we describe how SPAMDA deals with incomplete data when creating intermediate datasets and performing the matching process.

>>>>

%=================================================================
% Full title of the paper (Capitalized)
\Title{\textcolor{red}{A Software Tool for building Suitable Datasets for Softcomputing and Machine Learning Techniques from Meteorological Data Integration}}

@@

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{Meteorological data are extensively used to perform environmental learning. \textcolor{red}{Soft Computing (SC) and} Machine Learning (ML) techniques, a valuable support in many research areas, require datasets containing information related to the topic under study, which are not always available in an appropriate format, and its preparation and pre-processing implies a lot of time and effort by the researchers. This paper presents a novel software tool with an user-friendly GUI to create datasets by means of management and data integration of meteorological observations from two well-known sources of information: the \textit{National Data Buoy Center} and the \textit{National Centers for Environmental Prediction and for Atmospheric Research Reanalysis Project}. Such datasets can be created using buoys and reanalysis data by customisable procedures, in terms of predictive and objective variables and temporal resolution. These datasets can be used by \textcolor{red}{SC and} ML methodologies for prediction tasks (classification or regression) that support improvement of sustainability energy production, design of production systems as WECs or environmental modelling, among others.}

% Keywords
\keyword{Meteorological data; Reanalysis data; Environmental prediction; Marine energy; \textcolor{red}{Renewable Energy resource evaluation};  \textcolor{red}{Suitable Datasets for user models};}  % List three to ten pertinent keywords specific to the article, yet reasonably common within the subject discipline.

@@

		A better understanding of the environment is of vital importance for science, contributing not only to more efficient exploitation of natural resources but also to the development of new strategies aimed at its protection. In that sense, meteorological observations provide an essential and valuable source of information which is widely used by researchers to address environmental learning, comprehension, prediction and conservation in numerous oceanic and atmospheric studies of a wide variety of areas (e.g. energy, climate change, agriculture, etc.). Some specific examples of the diversity of fields in which meteorological data can be used in are, among others: estimation of global solar radiation based on sunshine duration \cite{SHAHRUKHANIS2019179}, directional analysis of sea storms \cite{LAFACE201545}, assessment and control of wind turbine \cite{DAHHANI2018167}, wind power ramp events prediction \cite{DORADOMORENO2017428}, evaluation of salinity-gradient energy harvesting \cite{REYESMENDOZA2020100645}, study of the responses exhibited by plankton to fluid motions \cite{FUCHS2016109}, trends in solar radiation \cite{SILVA20101852} or simulation of extreme near shore sea conditions \cite{GOULDBY201415} \textcolor{red}{Poner o sustituir aquí por alguna referencia de Energies}. All these studies require a prior data collection and its adaptation to a specific format that allows the interpretation of them.

		\textcolor{red}{Once good quality and well-formed data are obtained, these can be used to extract information and building prediction models that explain the behavior of a certain problem. Although Soft Computing (SC) and Machine Learning (ML) techniques have the ability to handle uncertainty in data and are extensively used for modeling purposes, the real challenge in modelling studies is due to the inadequacy of data, since the adequacy of the models depend mainly on the the quality of the information used, so that if a researcher does not have quality data, you will not obtain quality models.}

		\textcolor{red}{Continuing with this line}, special purpose software is usually developed to help researchers to advance in their studies related to energy and environmental modelling, becoming a great support for decision-making in the exploitation and protection of the environment. In \cite{CAMARGONOGUEIRA2016361}, a software tool for designing solar water heating systems is developed, which simulates different situations and finds the best technical and economical solution. In \cite{LO2015293}, an integrated simulation tool for the optimum design of bifacial solar panel with reflectors is presented. This tool can also be used to analyse the performance of the solar cells. A framework for data integration of offshore wind farms is implemented in \cite{NGUYEN2013150} in order to facilitate data exchange and improve operation and maintenance practices. In \cite{SHEHADEH2020100644}, a load management application to maximize the used energy of thermal systems and reduce the costs and emissions is developed. A risk assessment tool to improve safety standards and emergency management in onshore wind farms, is presented in \cite{ASTIASOGARCIA201648}. Raabe et al. \cite{RAABE2010213} developed two software tools, \textit{Model of Equilibrium of Bay Beaches} (MEPBAY) and \textit{Coastal Modelling System} (SMC), to support different operational levels of headland-bay beach in coastal engineering projects, and Motahhir et al. \cite{MOTAHHIR20199} developed an open hardware/software test bench for solar tracker.

@@

		Currently, and as a support to traditional study procedures, \textcolor{red}{SC and ML techniques} \cite{Alpaydin:2004:IML:1036287,Bishop:2006:PRM:1162264} are being widely used in numerous research fields related to classification, regression and optimisation tasks, obtaining significant improvements in the performance of the results \cite{DELFANI2019100559,NASRUDDIN201948,GUO201816} \textcolor{red}{(Sustituir algunas de las 5 referencias anteriores por algunas de Energies donde se apliquen tecnicas de SC y ML)}. \textcolor{red}{SC and} ML methodologies can be used not only by experienced computer scientists but also by other researchers. For example, the well-known \textit{Waikato Environment for Knowledge Analysis} (WEKA) \cite{WEKA} software tool provides researchers with a wide collection of ML algorithms. In this way, ML techniques have been already applied to tackle wave characterisation, accurately estimating $H_{s}$ and $T_{e}$ parameters \cite{DURANROSAL2017268, KUMAR2017605}, given that robustness of ML methods can tackle the previously explained difficulties in wave energy prediction. The problem is that, in order to apply these methods, it is essential to obtain datasets with relevant information about the issue under study, used to infer knowledge. Usually, these datasets are not publicly available in a friendly format, and their generation is the first step needed.

@@

		Up to now, and to the best of our knowledge, this is the first software tool addressing the problem previously discussed, integrating meteorological data from NDBC and NNRP. The novelties and functionalities that SPAMDA offers to the researchers will be detailed in Section \ref{sec:SPAMDA}, although some of them are briefly summarised below \textcolor{red}{Esto hay que revisarlo}:
		\begin{itemize}
			\item The generation of datasets becomes a very easy and customisable task, by means of the selection of different input parameters, such as predictive and objective variables, classification and regression ML techniques, output discretisation or prediction horizon, among others.
			\item It makes the researcher focus on environmental modelling, without having to worry about mechanical tasks, avoiding laborious pre-processing procedures that imply a lot of time and effort. \textcolor{red}{Decir que permite un estudio previo o sondeo antes de hacer estudios más exhaustivos}
			\item It manages the extensive casuistry of data integration which can lead to incomplete datasets.
			\item It avoids possible researcher errors in the intermediate steps of the process, such as time and geographical coordinates conversion or missing values handling (dates or measurements not recorded), among others.
			\item It provides information about the quality and quantity of the data.
			\item It includes different pre-processing tasks, such as normalisation and missing data recovery.\textcolor{red}{Hay que quitarle importancia al tema del preprocesamiento, para que no nos comparen con eso.}
			\item It facilitates data management and well-organised storage of the datasets.
			\item The created datasets can be easily used by ML tools.
			\item Its modular design allows the implementation of new modules for managing meteorological data from others sources, benefiting future renewable energy and environmental research.
			\item It includes a user-friendly GUI, facilitating and greatly simplifying data management, and it is integrated with the Explorer environment of WEKA.
			\item It is multi-platform, and it can be used on any computer with Java regardless of the operating system.
		\end{itemize}

@@

			Note that the data collected by the network of buoys may be incomplete due to diverse circumstances such as the weather conditions in which the buoys have to operate, failures or malfunctioning elements of the buoys, among others. Accordingly, it may be the situation that some of the measurements are completely missing (missing date or instance) or partially missing (some measurements not recorded), by a buoy or by a set of buoys, once in a while or over a period of time. It may be also possible that the measurements have been recorded at a time different from the expected one. These aspects have to be taken into account when creating the datasets. This casuistry is explained in detail in Appendix \ref{app:AppendixA}.

@@

				\textcolor{red}{Aquí en esta subsección hay que intentar darle poca importancia respecto a que el software no se centra en los filtros de preprocesamiento}

@@

				The reader can check in Appendix \ref{app:AppendixA} an example with a more complex case of the procedure.
				\begin{figure}[ht!]
					\centering
					\includegraphics[scale=0.38]{figures/FigureMatchingProcess.png}
					\caption{An example of matching the data from NDBC (left) and NNRP (right).}
					\label{fig:matchingProcess}
				\end{figure}

@@

\appendix
\section{Managing incomplete data}\label{app:AppendixA}

        \textcolor{red}{¿Como podriamos ponerle otro titulo para que no se asemeje a manual de usuario? Tenemos que mirar a ver si podemos tambien cambiar el titulo de alguna otra seccion en cuanto a ese fin.}

		In this appendix, we describe how SPAMDA deals with incomplete data when creating intermediate datasets and performing the matching process.
